Compilation and Backend-Independent Vectorization for Multi-Party Computation

### 2022-04-27 overleaf/paper_SIMD.tex
Edits to Section 3.

\begin{table*}
\begin{tabular}{ccc}
\begin{minipage}{0.275\textwidth}
{\small
\begin{pythonn}
def biometric(C: shared[list[int]], D: int, 
      S: shared[list[int]], N: int) -> 
      shared[tuple[int,int]]:
   min_sum = MAX_INT
   min_idx = 0  
   for i in range(N): 
      sum = 0
      for j in range(D): 
         # d = S[i,j] - C[j]
         d = S[i * D + j] - C[j] 
         p = d * d
         sum = sum + p
      if sum < min_sum:
         min_sum = sum
         min_idx = i
   return (min_sum, min_idx)
\end{pythonn}
}
\end{minipage}

@@

Source syntax is essentially standard IMP syntax as shown below:
%\begin{figure}[t]
{\small
\[
\begin{array}{l@{~~~~~}l}
  \begin{array}{l@{~}l@{~~~}l}
 e & ::= e \; \mathit{op} \; e \mid e \mid \x \mid  \texttt{const} \mid \texttt{A[$e$]} & \mathit{expression} \\
 % s &::= t \;\f & \mathit{field} \\
 %\md &::= t\;{\m}(t \;{\this}, \; t\;\x) \; \{\; \ol{t\;\y}\;\s;\;\return\;\y \; \} 
 %                 & \mathit{method} \\ 
  s & ::= s;s \mid  & \mathit{sequence} \\
  & \x = e \mid \texttt{A[$e$]} = e \mid & \mathit{assignment \; stmt} \\
  & \texttt{for $i$ in range($I$)}: s \mid & \mathit{for \; stmt} \\
  & \texttt{if $e$ then $s$ else $s$} & \mathit{if \; stmt} \\ 
%   t &::= q \; {\sf C} & \mathit{qualified \; type}\\
%   q &::= \high \mid \poly \mid \low & \mathit{FlowCFL \; qualifier}\\
%    t' &::= q' \; \C & \mathit{field \; qualified \; type}\\
%   q' &::= \tainted \mid \poly & \mathit{field \; qualifier}\\

@@

The type system is mostly standard, and in our experience, a sweet spot between simplicity and expressivity.
The \texttt{shared} qualifier denotes shared values, i.e., ones shared among the parties and computed upon 
under secure computation protocols vs. \texttt{plain} which denotes plaintext values. Shared lists denote 
shared elements, though the bounds of the list are plaintext. \ana{add subtyping}

Our compiler imposes certain semantic restrictions that it enforces throughout the various 
phased of compilation. We note that in some cases, the restrictions 
can be easily lifted and we plan to do so in future iterations of our compiler.

\begin{enumerate}
\item Loops are of the form \texttt{0 <= i < I} and bounds are fixed at compile time. 
It is a standard restriction in MPC that the bounds must be known at circuit-generation time.
\item Arrays are one-dimensional. N-dimensional arrays are linearized and accessed 
in row-major order and at this point the programmer is responsible for linearization
and access. (This restriction can be easily lifted.)
\item Array subscrpts are plaintext values. The typing rule for array access \texttt{A[$e$]} \ana{???}
Our compiler allows for output (write) arrays, however it restricts
write access to \emph{canonical writes} along the dimensions of the array. I.e., \texttt{A[i,j] = ...}
where \texttt{i} and \texttt{j} loop over the two dimensions of \texttt{A} is allowed, 
but \texttt{A[i,j+2] = ...} is not allowed. Read access can be an arbitrary function of the 
indices of the enclosing loops. 
\item \ana{Add rules for logical ops}
\item \ana{Add rules for arithmetic ops}
\item \ana{Add rules for MUX}
\end{enumerate}

We write $i,j,k$ to denote the loop nest: $i$ is the outermost loop, $j$, is immediately nested in $i$, and so on until $k$
and we use $I,J,K$ to denote the corresponding upper bounds. For simplicity, we write $A[i,j,k]$ to denote canonical access 
to an array element. In the program, canonical access is achieved via the standard row-major order formula: $(J*K)*i + K*j + k$.
To simplify the presentation we describe our algorithms in terms of three-element tuples $i,j,k$.
All discussion generalizes to arbitrarily large loop nests.

\subsection{Semantic Analysis}

@@

\subsection{MUX Nodes and Pseudo $\phi$-nodes}

>>>>

\begin{table*}
\begin{tabular}{ccc}
\begin{minipage}{0.275\textwidth}
{\small
\begin{pythonn}
def biometric(C: shared[list[int]], D: int, 
      S: shared[list[int]], N: int) -> 
      shared[tuple[int,int]]:
   min_sum : int = MAX_INT
   min_idx : int = 0  
   for i in range(N): 
      sum : int = 0
      for j in range(D): 
         # d = S[i,j] - C[j]
         d : int = S[i * D + j] - C[j] 
         p : int = d * d
         sum = sum + p
      if sum < min_sum:
         min_sum : int = sum
         min_idx : int = i
   return (min_sum, min_idx)
\end{pythonn}
}
\end{minipage}

@@

Source syntax is essentially standard IMP syntax as shown below:
%\begin{figure}[t]
{\small
\[
\begin{array}{l@{~~~~~}l}
  \begin{array}{l@{~}l@{~~~}l}
 e & ::= e \; \mathit{op} \; e \mid e \mid \x \mid  \texttt{const} \mid \texttt{A[$e$]} & \mathit{expression} \\
 % s &::= t \;\f & \mathit{field} \\
 %\md &::= t\;{\m}(t \;{\this}, \; t\;\x) \; \{\; \ol{t\;\y}\;\s;\;\return\;\y \; \} 
 %                 & \mathit{method} \\ 
  s & ::= s;s \mid  & \mathit{sequence} \\
  & \x = e \mid \texttt{A[$e$]} = e \mid & \mathit{assignment \; stmt} \\
  & \texttt{for $i$ in range($I$)}: s \mid & \mathit{for \; stmt} \\
  & \texttt{if $e$: $s$ else: $s$} & \mathit{if \; stmt} \\ 
%   t &::= q \; {\sf C} & \mathit{qualified \; type}\\
%   q &::= \high \mid \poly \mid \low & \mathit{FlowCFL \; qualifier}\\
%    t' &::= q' \; \C & \mathit{field \; qualified \; type}\\
%   q' &::= \tainted \mid \poly & \mathit{field \; qualifier}\\

@@

The type system is mostly standard, and in our experience, a sweet spot between readability and expressivity.
The \texttt{shared} qualifier denotes shared values, i.e., ones shared among the parties and computed upon 
under secure computation protocols; the \texttt{plain} qualifier denotes plaintext values. Shared lists denote 
shared elements, but plaintext length. Subtyping is again as expected: $q_1 \; \tau <: q_2 \; \tau$ 
iff $q_1 <: q_2$ where $\texttt{plain} <: \texttt{shared}$. Thus, we can convert a plaintext value into a shared one, 
but not vice versa.

Our compiler imposes certain semantic restrictions that it enforces throughout the various 
phases of compilation. We note that in some cases, the restrictions 
can be easily lifted and we plan to do so in future iterations of the work.

\begin{enumerate}
\item Loops are of the form $0 <= i < I$ and bounds are fixed at compile time. 
It is a standard restriction in MPC that the bounds must be known at circuit-generation time.
\item Arrays are one-dimensional. N-dimensional arrays are linearized and accessed 
in row-major order and at this point the programmer is responsible for linearization
and access. (This restriction can be easily lifted.)
\item Array subscrpts are plaintext values as specified by the rule:
\begin{semantics}
    \ntyperule{Array Access}{
%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
  \Gamma \vdash e : \texttt{plain int} \quad \Gamma \vdash \texttt{A} : q \; \texttt{list}[\tau] \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
}{
  \Gamma \vdash \texttt{A[$e$]} : q \; \tau
}
\end{semantics}
The subscript $e$ is a function of the indices of the enclosing loops. 
For read access, the compiler allows an arbitrary such function. 
However, it restricts write access to \emph{canonical writes}, i.e., \texttt{A[i,j,k]} = ...
where \texttt{i}, \texttt{j} and \texttt{k} loop over the three dimensions of \texttt{A}.  
Write access such as for example \texttt{A[i,j+2]} = ... is not allowed. 
%\item \ana{Add rules for logical ops}
%\item \ana{Add rules for arithmetic ops}
\item The final restriction involves MUX as expressed by the rule: 
\begin{semantics}
    \ntyperule{MUX}{
%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
  \Gamma \vdash e_1 : q_1 \; \texttt{bool} \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
}{
  \Gamma \vdash \texttt{MUX($e_1$,$e_2$,$e_3$)} : q_1 \vee q_2 \vee q_3 \; \tau
}
\end{semantics}
\end{enumerate}
Importantly, the arguments of MUX are restricted to base types, as passing array arguments 
is not valid MOTION MPC. \ana{Check this.} This causes a slight inconvenience as we could not write 
\begin{pythonn}
if e: A[i] = val
\end{pythonn}
Instead we had to write
\begin{pythonn}
if not(e): val = A[i];
A[i] = val;
\end{pythonn}

We note that while the programmer writes 
annotations at the level of IMP Source (recall \ana{ref to overview section}), the 
annotations propagate through the transformations and are checked at the 
level of MPC Source.

%\ana{Disadvantage of MUX}

For the rest of this section we write $i,j,k$ to denote the loop nest: $i$ is the outermost loop, $j$, is immediately nested in $i$, and so on until $k$
and we use $I,J,K$ to denote the corresponding upper bounds. We write $A[i,j,k]$ to denote canonical access 
to an array element. In the program, canonical access is achieved via the standard row-major order formula: $(J*K)*i + K*j + k$.
To simplify the presentation we describe our algorithms in terms of three-element tuples $i,j,k$, however, discussion easily generalizes to 
arbitrarily large loop nests.

\subsection{From IMP Source to SSA}

@@

\subsection{From SSA to MPC Source: MUX Nodes and Pseudo $\phi$-nodes}
